---
title: "Gene Classification (adjusted from MG)"
author: "Brooke Peckenpaugh"
date: "6 Dec 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~/Desktop/data/aim_3")

library(tidyverse)
```


### Functions
```{r}
classify_expressed <- function(row, names, cutoff){
  #Checks if gene is expressed greater than [cutoff] in at least 1 of the tissues included in  names.
  #Returns boolean
  
  row <- row[which(names(row) %in% names)]
  for(i in seq(1, length(names))){
    if (as.numeric(row[i]) > cutoff){
      return(TRUE)
    }
  }
  return(FALSE)
}

classify_trace <- function(row, names, cutoff){
  #Checks if gene is expressed less than [cutoff] in ALL of the tissues included in names.
  #Returns boolean
  
  num_trace <-  0
  
  row <- row[which(names(row) %in% names)]
  for(i in seq(1, length(names))){
    if (as.numeric(row[i]) < cutoff){
      num_trace <- num_trace + 1
    }
  }
  if  (num_trace == length(names)){
    return(TRUE)
  }else{return(FALSE)}
  
}

#calculate tau measure of tissue specificity
#https://github.com/severinEvo/gene_expression
tau<-function(x){
  if(any(is.na(x))) stop('NA\'s need to be 0.')
  if(any(x<0)) stop('Negative input values not permitted.')
  t<-sum(1-x/max(x))/(length(x)-1)
}

```


# Load  data
## melanogaster (ore)
```{r}
mel <- read.table("~/Desktop/data/aim_3/dmel.transTPM.txt", header = T)

mel_FB <- read.table("~/Desktop/data/aim_3/GSE99574_HiSAT2_dmel.transcript_level_TPM.FB.txt", header = T)

mel_ore_f_full <- mel %>% select(1, 3, starts_with("oreR_f"))

mel_ore_f <- mel_ore_f_full %>% 
  mutate(mean_head = rowMeans(select(mel_ore_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(mel_ore_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(mel_ore_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(mel_ore_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(mel_ore_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(mel_ore_f_full, contains("_re_"))),
        mean_genitalia = rowMeans(select(mel_ore_f_full, contains("_ge_")))) %>%
  select(1, 2, starts_with("mean"))


mel_ore_m_full <- mel %>% select(1, 3, starts_with("oreR_m"))

mel_ore_m <- mel_ore_m_full %>% 
  mutate(mean_head = rowMeans(select(mel_ore_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(mel_ore_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(mel_ore_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(mel_ore_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(mel_ore_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(mel_ore_m_full, contains("_re_"))),
        mean_genitalia = rowMeans(select(mel_ore_m_full, contains("_ge_")))) %>%
  select(1, 2, starts_with("mean"))

```

## melanogaster (w1118)
```{r}
mel <- read.table("~/Desktop/data/aim_3/dmel.transTPM.txt", header = T)

mel_w1118_f_full <- mel %>% select(1, starts_with("w1118_f"))

mel_w1118_f <- mel_w1118_f_full %>% 
  mutate(mean_head = rowMeans(select(mel_w1118_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(mel_w1118_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(mel_w1118_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(mel_w1118_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(mel_w1118_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(mel_w1118_f_full, contains("_re_"))),
        mean_genitalia = rowMeans(select(mel_w1118_f_full, contains("_ge_")))) %>%
  select(1, starts_with("mean"))


mel_w1118_m_full <- mel %>% select(1, starts_with("w1118_m"))

mel_w1118_m <- mel_w1118_m_full %>% 
  mutate(mean_head = rowMeans(select(mel_w1118_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(mel_w1118_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(mel_w1118_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(mel_w1118_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(mel_w1118_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(mel_w1118_m_full, contains("_re_"))),
        mean_genitalia = rowMeans(select(mel_w1118_m_full, contains("_ge_")))) %>%
  select(1, starts_with("mean"))

```

## ananassae
```{r}
ana <- read.table("~/Desktop/data/aim_3/dana.transTPM.txt", header = T, fill = T)
ana <- na.omit(ana)

ana_f_full <- ana %>% select(1, starts_with("dana_f"))

ana_f <- ana_f_full %>% 
  mutate(mean_head = rowMeans(select(ana_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(ana_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(ana_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(ana_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(ana_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(ana_f_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))


ana_m_full <- ana %>% select(1, starts_with("dana_m"))

ana_m <- ana_m_full %>% 
  mutate(mean_head = rowMeans(select(ana_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(ana_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(ana_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(ana_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(ana_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(ana_m_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))

```

## mojavensis
```{r}
moj <- read.table("~/Desktop/data/aim_3/dmoj.transTPM.txt", header = T, fill = T)
moj <- na.omit(moj)

moj_f_full <- moj %>% select(1, starts_with("dmoj_f"))

moj_f <- moj_f_full %>% 
  mutate(mean_head = rowMeans(select(moj_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(moj_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(moj_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(moj_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(moj_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(moj_f_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))


moj_m_full <- moj %>% select(1, starts_with("dmoj_m"))

moj_m <- moj_m_full %>% 
  mutate(mean_head = rowMeans(select(moj_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(moj_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(moj_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(moj_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(moj_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(moj_m_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))

```

## persimilis
```{r}
per <- read.table("~/Desktop/data/aim_3/dper.transTPM.txt", header = T, fill = T)
per <- na.omit(per)

per_f_full <- per %>% select(1, starts_with("dper_f"))

per_f <- per_f_full %>% 
  mutate(mean_head = rowMeans(select(per_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(per_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(per_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(per_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(per_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(per_f_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))


per_m_full <- per %>% select(1, starts_with("dper_m"))

per_m <- per_m_full %>% 
  mutate(mean_head = rowMeans(select(per_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(per_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(per_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(per_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(per_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(per_m_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))

```

## pseudoobscura
```{r}
pse <- read.table("~/Desktop/data/aim_3/dpse.transTPM.txt", header = T, fill = T)
pse <- na.omit(pse)

pse_f_full <- pse %>% select(1, starts_with("dpse_f"))

pse_f <- pse_f_full %>% 
  mutate(mean_head = rowMeans(select(pse_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(pse_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(pse_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(pse_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(pse_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(pse_f_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))


pse_m_full <- pse %>% select(1, starts_with("dpse_m"))

pse_m <- pse_m_full %>% 
  mutate(mean_head = rowMeans(select(pse_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(pse_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(pse_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(pse_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(pse_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(pse_m_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))

```

## virilis
```{r}
vir <- read.table("~/Desktop/data/aim_3/dvir.transTPM.txt", header = T, fill = T)
vir <- na.omit(vir)

vir_f_full <- vir %>% select(1, starts_with("dvir_f"))

vir_f <- vir_f_full %>% 
  mutate(mean_head = rowMeans(select(vir_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(vir_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(vir_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(vir_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(vir_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(vir_f_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))


vir_m_full <- vir %>% select(1, starts_with("dvir_m"))

vir_m <- vir_m_full %>% 
  mutate(mean_head = rowMeans(select(vir_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(vir_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(vir_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(vir_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(vir_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(vir_m_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))

```

## willistoni
```{r}
wil <- read.table("~/Desktop/data/aim_3/dwil.transTPM.txt", header = T, fill = T)
wil <- na.omit(wil)

wil_f_full <- wil %>% select(1, starts_with("dwil_f"))

wil_f <- wil_f_full %>% 
  mutate(mean_head = rowMeans(select(wil_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(wil_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(wil_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(wil_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(wil_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(wil_f_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))


wil_m_full <- wil %>% select(1, starts_with("dwil_m"))

wil_m <- wil_m_full %>% 
  mutate(mean_head = rowMeans(select(wil_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(wil_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(wil_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(wil_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(wil_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(wil_m_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))

```

## yakuba
```{r}
yak <- read.table("~/Desktop/data/aim_3/dyak.transTPM.txt", header = T, fill = T)
yak <- na.omit(yak)

yak_f_full <- yak %>% select(1, starts_with("dyak_f"))

yak_f <- yak_f_full %>% 
  mutate(mean_head = rowMeans(select(yak_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(yak_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(yak_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(yak_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(yak_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(yak_f_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))


yak_m_full <- yak %>% select(1, starts_with("dyak_m"))

yak_m <- yak_m_full %>% 
  mutate(mean_head = rowMeans(select(yak_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(yak_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(yak_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(yak_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(yak_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(yak_m_full, contains("_re_")))) %>%
  select(1, starts_with("mean"))

```

## grimshawi
```{r}
haw <- read.table("~/Desktop/data/aim_3/haw.transTPM.txt", header = T, fill = T)
haw <- na.omit(haw)

gri_f_full <- haw %>% select(1, starts_with("dgriG1_f"))

gri_f <- gri_f_full %>% 
  mutate(mean_head = rowMeans(select(gri_f_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(gri_f_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(gri_f_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(gri_f_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(gri_f_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(gri_f_full, contains("_re_"))),
        mean_genitalia = rowMeans(select(gri_f_full, contains("_ge_")))) %>%
  select(1, starts_with("mean"))


gri_m_full <- haw %>% select(1, starts_with("dgriG1_m"))

gri_m <- gri_m_full %>% 
  mutate(mean_head = rowMeans(select(gri_m_full, contains("_hd_"))),
        mean_thorax = rowMeans(select(gri_m_full, contains("_tx_"))), 
        mean_abdomen = rowMeans(select(gri_m_full, contains("_ac_"))),
        mean_digestive = rowMeans(select(gri_m_full, contains("_dg_"))),
        mean_gonad = rowMeans(select(gri_m_full, contains("_go_"))),
        mean_reproductive = rowMeans(select(gri_m_full, contains("_re_"))),
        mean_genitalia = rowMeans(select(gri_m_full, contains("_ge_")))) %>%
  select(1, starts_with("mean"))

```


# Calculate Tau

## mel (ore) female
```{r}
mel_ore_f_NA <- replace(mel_ore_f, mel_ore_f < 1, NA)
mel_ore_f_NA_log <- log2(mel_ore_f_NA[,3:9])
mel_ore_f_log <- replace(mel_ore_f_NA_log, is.na(mel_ore_f_NA_log), 0)

mel_ore_f_log$tau <- apply(mel_ore_f_log, 1, tau)

mel_ore_f_tau <- cbind(mel_ore_f[,1:2], mel_ore_f_log)

#filter rows where reproductive tissues have highest gene expression
mel_ore_f_RP <- mel_ore_f_tau %>% 
  filter(mean_gonad > mean_head & mean_gonad > mean_thorax & mean_gonad > mean_abdomen & mean_gonad > mean_digestive |
           mean_reproductive > mean_head & mean_reproductive > mean_thorax & mean_reproductive > mean_abdomen & mean_reproductive > mean_digestive |
           mean_genitalia > mean_head & mean_genitalia > mean_thorax & mean_genitalia > mean_abdomen & mean_genitalia > mean_digestive)

#filter tissue-specific, reproductive genes
mel_ore_f_RPspec <- mel_ore_f_RP %>% filter(tau > 0.8)


```

## mel (ore) male
```{r}
mel_ore_m_NA <- replace(mel_ore_m, mel_ore_m < 1, NA)
mel_ore_m_NA_log <- log2(mel_ore_m_NA[,3:9])
mel_ore_m_log <- replace(mel_ore_m_NA_log, is.na(mel_ore_m_NA_log), 0)

mel_ore_m_log$tau <- apply(mel_ore_m_log, 1, tau)

mel_ore_m_tau <- cbind(mel_ore_m[,1:2], mel_ore_m_log)

#filter rows where reproductive tissues have highest gene expression
mel_ore_m_RP <- mel_ore_m_tau %>% 
  filter(mean_gonad > mean_head & mean_gonad > mean_thorax & mean_gonad > mean_abdomen & mean_gonad > mean_digestive |
           mean_reproductive > mean_head & mean_reproductive > mean_thorax & mean_reproductive > mean_abdomen & mean_reproductive > mean_digestive |
           mean_genitalia > mean_head & mean_genitalia > mean_thorax & mean_genitalia > mean_abdomen & mean_genitalia > mean_digestive)

#filter tissue-specific, reproductive genes
mel_ore_m_RPspec <- mel_ore_m_RP %>% filter(tau > 0.8)


```














# RP, SM, & GR

## Classify

### Overview of what I'm doing to classify genes:

For each species...
1. Identify SomExpressed genes (> 2 TPM in at least one somatic tissue; TRUE/FALSE)
2. Identify RepExpressed genes (> 2 TPM in at least one reproductive tissue; TRUE/FALSE)
3. Identify somatic trace expression genes (< 0.5 TPM in all som tissues; TRUE/FALSE)
4. Identify reproductive trace expression genes (< 0.5 TPM in all rep tissues; TRUE/FALSE)

5. Classify GR genes as those which are SomExpressed AND RepExpressed AND not expressed at trace levels in either (TRUE/FALSE)
6. Classify SM genes as those which are SomExpressed AND trace expressed in reproductive tissues AND not in the GR list (TRUE/FALSE)
7. Classify RP genes as those which are RepExpressed AND trace expressed in somatic tissues AND not in the GR list (TRUE/FALSE)


## MELANOGASTER

```{r}
upperCutoff <- 2
lowerCutoff <- 0.5

```

### mel (ore) female
```{r}
mel_ore_f <- mel_ore_f %>% mutate(SomExpressed = apply(mel_ore_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(mel_ore_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), upperCutoff),
                      SomTrace = apply(mel_ore_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(mel_ore_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


mel_ore_f_RP <- mel_ore_f %>% filter(RP == TRUE)
mel_ore_f_SM <- mel_ore_f %>% filter(SM == TRUE)
mel_ore_f_GR <- mel_ore_f %>% filter(GR == TRUE)

```

### mel (ore) male
```{r}
mel_ore_m <- mel_ore_m %>% mutate(SomExpressed = apply(mel_ore_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(mel_ore_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), upperCutoff),
                      SomTrace = apply(mel_ore_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(mel_ore_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


mel_ore_m_RP <- mel_ore_m %>% filter(RP == TRUE)
mel_ore_m_SM <- mel_ore_m %>% filter(SM == TRUE)
mel_ore_m_GR <- mel_ore_m %>% filter(GR == TRUE)

```

### mel (w1118) female
```{r}
mel_w1118_f <- mel_w1118_f %>% mutate(SomExpressed = apply(mel_w1118_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(mel_w1118_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), upperCutoff),
                      SomTrace = apply(mel_w1118_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(mel_w1118_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


mel_w1118_f_RP <- mel_w1118_f %>% filter(RP == TRUE)
mel_w1118_f_SM <- mel_w1118_f %>% filter(SM == TRUE)
mel_w1118_f_GR <- mel_w1118_f %>% filter(GR == TRUE)

```

### mel (w1118) male
```{r}
mel_w1118_m <- mel_w1118_m %>% mutate(SomExpressed = apply(mel_w1118_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(mel_w1118_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), upperCutoff),
                      SomTrace = apply(mel_w1118_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(mel_w1118_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


mel_w1118_m_RP <- mel_w1118_m %>% filter(RP == TRUE)
mel_w1118_m_SM <- mel_w1118_m %>% filter(SM == TRUE)
mel_w1118_m_GR <- mel_w1118_m %>% filter(GR == TRUE)

```


## ANANASSAE

### ana female
```{r}
ana_f <- ana_f %>% mutate(SomExpressed = apply(ana_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen", "mean_digestive"), upperCutoff),
                      RepExpressed = apply(ana_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(ana_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(ana_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


ana_f_RP <- ana_f %>% filter(RP == TRUE)
ana_f_SM <- ana_f %>% filter(SM == TRUE)
ana_f_GR <- ana_f %>% filter(GR == TRUE)

```

### ana male
```{r}
ana_m <- ana_m %>% mutate(SomExpressed = apply(ana_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(ana_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(ana_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(ana_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


ana_m_RP <- ana_m %>% filter(RP == TRUE)
ana_m_SM <- ana_m %>% filter(SM == TRUE)
ana_m_GR <- ana_m %>% filter(GR == TRUE)

```


## MOJAVENSIS

### moj female
```{r}
moj_f <- moj_f %>% mutate(SomExpressed = apply(moj_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(moj_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(moj_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(moj_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


moj_f_RP <- moj_f %>% filter(RP == TRUE)
moj_f_SM <- moj_f %>% filter(SM == TRUE)
moj_f_GR <- moj_f %>% filter(GR == TRUE)

```

### moj male
```{r}
moj_m <- moj_m %>% mutate(SomExpressed = apply(moj_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(moj_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(moj_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(moj_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


moj_m_RP <- moj_m %>% filter(RP == TRUE)
moj_m_SM <- moj_m %>% filter(SM == TRUE)
moj_m_GR <- moj_m %>% filter(GR == TRUE)

```


## PERSIMILIS

### per female
```{r}
per_f <- per_f %>% mutate(SomExpressed = apply(per_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(per_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(per_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(per_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


per_f_RP <- per_f %>% filter(RP == TRUE)
per_f_SM <- per_f %>% filter(SM == TRUE)
per_f_GR <- per_f %>% filter(GR == TRUE)

```

### per male
```{r}
per_m <- per_m %>% mutate(SomExpressed = apply(per_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(per_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(per_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(per_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


per_m_RP <- per_m %>% filter(RP == TRUE)
per_m_SM <- per_m %>% filter(SM == TRUE)
per_m_GR <- per_m %>% filter(GR == TRUE)

```


## PSEUDOOBSCURA

### pse female
```{r}
pse_f <- pse_f %>% mutate(SomExpressed = apply(pse_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(pse_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(pse_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(pse_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


pse_f_RP <- pse_f %>% filter(RP == TRUE)
pse_f_SM <- pse_f %>% filter(SM == TRUE)
pse_f_GR <- pse_f %>% filter(GR == TRUE)

```

### pse male
```{r}
pse_m <- pse_m %>% mutate(SomExpressed = apply(pse_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(pse_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(pse_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(pse_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


pse_m_RP <- pse_m %>% filter(RP == TRUE)
pse_m_SM <- pse_m %>% filter(SM == TRUE)
pse_m_GR <- pse_m %>% filter(GR == TRUE)

```


## VIRILIS

### vir female
```{r}
vir_f <- vir_f %>% mutate(SomExpressed = apply(vir_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(vir_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(vir_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(vir_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


vir_f_RP <- vir_f %>% filter(RP == TRUE)
vir_f_SM <- vir_f %>% filter(SM == TRUE)
vir_f_GR <- vir_f %>% filter(GR == TRUE)

```

### vir male
```{r}
vir_m <- vir_m %>% mutate(SomExpressed = apply(vir_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(vir_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(vir_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(vir_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


vir_m_RP <- vir_m %>% filter(RP == TRUE)
vir_m_SM <- vir_m %>% filter(SM == TRUE)
vir_m_GR <- vir_m %>% filter(GR == TRUE)

```


## WILLISTONI

### wil female
```{r}
wil_f <- wil_f %>% mutate(SomExpressed = apply(wil_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(wil_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(wil_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(wil_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


wil_f_RP <- wil_f %>% filter(RP == TRUE)
wil_f_SM <- wil_f %>% filter(SM == TRUE)
wil_f_GR <- wil_f %>% filter(GR == TRUE)

```

### wil male
```{r}
wil_m <- wil_m %>% mutate(SomExpressed = apply(wil_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(wil_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(wil_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(wil_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


wil_m_RP <- wil_m %>% filter(RP == TRUE)
wil_m_SM <- wil_m %>% filter(SM == TRUE)
wil_m_GR <- wil_m %>% filter(GR == TRUE)

```


## YAKUBA

### yak female
```{r}
yak_f <- yak_f %>% mutate(SomExpressed = apply(yak_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(yak_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(yak_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(yak_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


yak_f_RP <- yak_f %>% filter(RP == TRUE)
yak_f_SM <- yak_f %>% filter(SM == TRUE)
yak_f_GR <- yak_f %>% filter(GR == TRUE)

```

### yak male
```{r}
yak_m <- yak_m %>% mutate(SomExpressed = apply(yak_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(yak_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive"), upperCutoff),
                      SomTrace = apply(yak_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(yak_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


yak_m_RP <- yak_m %>% filter(RP == TRUE)
yak_m_SM <- yak_m %>% filter(SM == TRUE)
yak_m_GR <- yak_m %>% filter(GR == TRUE)

```


## GRIMSHAWI

### gri female
```{r}
gri_f <- gri_f %>% mutate(SomExpressed = apply(gri_f, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(gri_f, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), upperCutoff),
                      SomTrace = apply(gri_f, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(gri_f, 1, classify_trace, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


gri_f_RP <- gri_f %>% filter(RP == TRUE)
gri_f_SM <- gri_f %>% filter(SM == TRUE)
gri_f_GR <- gri_f %>% filter(GR == TRUE)

```

### gri male
```{r}
gri_m <- gri_m %>% mutate(SomExpressed = apply(gri_m, 1, classify_expressed, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), upperCutoff),
                      RepExpressed = apply(gri_m, 1, classify_expressed, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), upperCutoff),
                      SomTrace = apply(gri_m, 1, classify_trace, names = c("mean_head", "mean_thorax", "mean_abdomen",  "mean_digestive"), lowerCutoff),
                      RepTrace = apply(gri_m, 1, classify_trace, names = c("mean_gonad", "mean_reproductive", "mean_genitalia"), lowerCutoff)) %>%
  mutate(GR = SomExpressed & RepExpressed & !SomTrace & !RepTrace,
         SM = SomExpressed & RepTrace & !GR,
         RP = RepExpressed & SomTrace & !GR)


gri_m_RP <- gri_m %>% filter(RP == TRUE)
gri_m_SM <- gri_m %>% filter(SM == TRUE)
gri_m_GR <- gri_m %>% filter(GR == TRUE)

```
